Some thought experiments on the applicability of Granger causality and Directed Information in inferring direction of information flows

### 2015-07-06 paper.tex
Add section of possible objections.

A review of a previous submission raised several objections to this work, which we seek to clarify here. Further, there are a few shortcomings to our analysis, which could be room for future work.

\begin{enumerate}
	\item Granger causality vs. true causality - cite Pearl. Our argument is, in spirit, very similar to the common criticism of Granger causality, which states that GC is a statistical quantity. It does not answer the question of true causal influence: if X is turned off, will Y behave differenty? (Need to quote from Pearl here).However, in the situation we are studying, we ask a different question - we wish to understand whether or not GC gives the correct direction of *information flow*. In our analysis, causal influences are present in both directions.
	\item Circularity - there is no circularity of argument. We are just assuming that there are sufficient number of trials to fully estimate the coefficients of the non-stationary AR process. We describe how this might be achieved in section ... (maybe this should come at the end?)
	\item Math is not complicated - this is a strength of our analysis. Even simple counter-examples that do not employ difficult proof techniques are able to show that GC does not always correctly estimate the direction of info flow.
	\item Not evolving source - not a stationary AR process. To circumvent this, we discuss estimation of non-stationary AR coeffs, since this is a neuroscience experiment, so we have many realizations of the same AR process in the form of multiple trials (cite Ding again. Quote if necessary) A future work will address this shortcoming by examining how GC works with an evolving source that produces a stationary process.
	\item We restrict ourselves to Gaussian noise for simplicity of analysis. Noise in the actual brain is likely to be very different. However, AR processes are often used to model ERPs (need a citation for this).
	\item In the S\&K algo, we have simplified power and energy constraints
	\item This is for a simple two-node network. But for larger networks, the issue of incorrectly estimating the direction will probably only get exacerbated.
\end{enumerate}

>>>>

A review of a previous submission raised some objections to this work, which we seek to rectify here. Further, our analysis has certain shortcomings, which we acknowledge here, and could be room for future work.

\begin{enumerate}
	\item Firstly, and most importantly, previous work has already shown that Granger Causality is not true causality (as defined by Judea Pearl~\cite{PearlCausalityBook}. Our argument is, in spirit, very similar to this common criticism of Granger causality, which states that Granger Causal influence is a statistical quantity and that it does not answer the question of true causal influence: if X is turned off, will Y behave differenty? {\color{red} (Need to quote from Pearl here)}.However, in the situation we are studying, we ask a different question - we wish to understand whether or not Granger Causality gives the correct direction of \emph{information flow}. In our analysis, causal influences are present in both directions, even in a ``true-causality'' sense.
	\item It might appear that we make a circular argument while computing the Granger Causal influences in our counter-example, since we supply the model for the stochastic process and use the theoretical coeffiecients of the model directly to compute the Granger Causality metric. However, as we state in Section~\ref{sec:figure-it-out}, we assume that the regression coefficients of the Autoregressive model can be exactly estimated (even if the AR process is non-stationary), and discuss how this might be achieved with the help of multiple trials. {\color{red} (Move to conclusions?)}
	\item Our analysis does not tackle an information source that evolves with time. Hence, our communication process (inspired from the scheme of Schalkwijk and Kailath~\cite{S&K}) is non-ergodic. Since Granger causality is really just relative errors in prediction of a process, in presence or absence of knowledge of another process, we compute the obvious generalization of Granger causality to non-ergodic processes. Nevertheless, future work will address a situation with a linear dynamical system as the information source.
	\item Our experiments restrict themselves to Gaussian noise for simplicity, but neural spiking and spike-rate models for spikes tend to be very different from those used here. This is a clear direction for future work.
	\item The power and energy constraints are somewhat oversimplified to make the analysis simpler. This is for simplicity of exposition. A more general analysis is a trivial extension.
	\item We consider a simple point-to-point network. In general networks, this issue could be even more complex. However, since this issue shows up even for the simplest network, we feel that the problem will only be exacerbated when the network is large.
	\item We would also like to note that this paper is \emph{not} mathematically deep. We believe that this is a strength of our analysis. Even simple counter-examples that do not employ difficult proof techniques are able to show that Granger Causality does not always correctly estimate the direction of information flow. {\color{red} (Move this?)}
\end{enumerate}
